Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-05-09 VegOzone.tex
Rephrase gpp co2 uptake sentence

The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a similar pattern as the generallized northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b). The spring peak occurs slightly earlier ($100\,\unit{doy}$) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology. This is in line with a general decrease in tropospheric ozone with decreasing altitude. The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013} compared to an estimated average begin of greening ($\left<G_\mathrm{start}\right>$) on $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details). In July--September, ozone is occasionally almost completely depleted during nighttime. This hint to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air. The derived \chem{\left<[O_3]\right>} is likely biased towards lower \chem{[O_3]} due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s. To account for this, we derive a bias correction factor by cross-calibrating with the timeseries taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).

>>>>

The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a similar pattern as the generallized northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b). The spring peak occurs slightly earlier ($100\,\unit{doy}$) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology. This is in line with a general decrease in tropospheric ozone with decreasing altitude. The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees has been observed as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013} compared to an estimated average begin of greening ($\left<G_\mathrm{start}\right>$) on $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details). In July--September, ozone is occasionally almost completely depleted during nighttime. This hint to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air. The derived \chem{\left<[O_3]\right>} is likely biased towards lower \chem{[O_3]} due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s. To account for this, we derive a bias correction factor by cross-calibrating with the timeseries taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).
